Project Description: 

STEP 1: Environment set up
1.1 Load required packages
1.2 set up work directory (based on where you save the file)
1.3 use fread to load 4 data tables
1.4 sanity check for data tables
```{r}
# 1
library(data.table)
library(sqldf)
library(doMC)
library(ROCR)
library(glmnet)
library(caret)
options(scipen=999)

# 2
path ='/Users/David/Desktop/Project\ Ecommece'
setwd(path)
# 3
customer_table <- fread("customer_table.csv")
order_table <- fread("order_table.csv")
product_table <- fread("product_table.csv")
category_table <- fread("category_table.csv")

# to change from scientific notation to the actual number
customer_table[,customer_id := as.character(customer_table$customer_id)]
order_table[,customer_id := as.character(order_table$customer_id)]
order_table[,order_id := as.character(order_table$order_id)]
order_table[,product_id := as.character(order_table$product_id)]
product_table[,product_id := as.character(product_table$product_id)]
```
if you meets error, try:  install.packages("the library name")


STEP 2: Get targeted audience
2.1  select users who didn't purchase anything between 2016/12/22 and 2017/02/22, aka this group of users went dormant for 3 months.
2.2 
2.3


```{r}
head(order_table)
attach(order_table)
have_purchase=sqldf(
  "SELECT customer_id, max(order_date) as Latest_order_time
  FROM order_table
  WHERE order_date between 20161222 and 20170222 and order_amount>0
  GROUP BY 1"
)

## find customers who were dormant between 2016/12/22 and 2017/02/22
dormant_3month = sqldf("SELECT * 
                        FROM base 
                        WHERE customer_id NOT IN 
                        (SELECT customer_id FROM purchase_again)")


## find customers who purchased again between 2017/02/23 and 2017/05/22
purchase_flag = sqldf("SELECT customer_id
                       FROM order_table
                       WHERE order_date BETWEEN '20170223' and '20170522'
                       GROUP BY 1")

is_converted = sqldf("SELECT CASE WHEN pf.customer_id IS NOT NULL THEN 1
                      ELSE 0 END AS is_converted 
                      FROM dormant_3month d
                      LEFT JOIN purchase_flag pf
                      ON d.customer_id = pf.customer_id
                      ")
## pull all features for the targeted audience
user_cohort <- sqldf("SELECT * 
                     FROM customer_table
                     WHERE customer_id IN (SELECT customer_id FROM dormant_3month)")

## combine the dependent variable with independent variables
user_features <- data.table(
cbind(is_converted,user_cohort,dormant_3month[,c("order_date","order_amount","product_id")]))

## remove temporary tables
rm(base,dormant_3month,is_converted,purchase_again,purchase_flag,user_cohort)
gc()

```

3. Data pre-processing
3.1 remove ML unrelated field

```{r}
## remove ML unrelated field
user_features[,c("customer_id","last_visit_date","age") := NULL]

## calculate new features, how long has this customer been around?
user_features[,tenure := as.numeric(as.Date("2017-02-22") - 
                                      as.Date(first_visit_date))]
user_features[,days_since_first_order := as.numeric(as.Date("2017-02-22") -
                                                      as.Date(as.character(order_date),format='%Y%m%d'))]



user_features <- data.table(
  sqldf(" 
        SELECT uf.*
        , CASE WHEN pt.category_id IS NOT NULL THEN pt.category_id ELSE 'unknown' END AS category_id
        FROM user_features uf
        LEFT JOIN product_table pt
        ON uf.product_id = pt.product_id;"))

user_features[,c("first_visit_date","order_date","product_id") := NULL]

## deal with missing value: fill all NA's with 0
user_features$category_id[is.na(user_features$category_id)] <- 'unknown'
user_features$latest_device_class[is.na(user_features$latest_device_class)] <- 'unknown'
user_features[is.na(user_features)] <- 0

## find out top 20 countries
top_countries <- as.character(as.matrix(user_features[,list(Count=.N), by = country][order(-Count)][1:30][,.(country)]))

user_features[,country_updated := ifelse(country %in% top_countries,country,'others')]
user_features[,c("country") := NULL]

## dummy variables
country_dummy <- model.matrix( ~ country_updated - 1, data=user_features)
gender_dummy <- model.matrix( ~ gender - 1, data=user_features)
category_dummy <- model.matrix( ~ category_id - 1, data=user_features)
deviceclass_dummy <- model.matrix( ~ latest_device_class - 1, data=user_features)

user_features_combined <- cbind(user_features,country_dummy,gender_dummy,category_dummy,deviceclass_dummy)
user_features_combined[,c("country_updated","gender","category_id","latest_device_class") := NULL]
rm(country_dummy,gender_dummy,category_dummy,deviceclass_dummy)

## zero variance variable
zero_variance <- names(Filter(function(x)(length(unique(x)) == 1),user_features_combined))
## filter out zero variance variable
user_features_combined <- Filter(function(x)(length(unique(x)) > 1),user_features_combined)

## transform continuous variables
transform_columns <- c("user_feature","phone_feature","tablet_feature","family_size","number_of_devices","tenure","order_amount","days_since_first_order")
transformed_column     <- user_features_combined[ ,grepl(paste(transform_columns, collapse = "|"),names(user_features_combined)),with = FALSE]
non_transformed_column <- user_features_combined[ ,-grepl(paste(transform_columns, collapse = "|"),names(user_features_combined)),with = FALSE]

transformed_column_processed <- predict(preProcess(transformed_column, method = c("BoxCox","scale")),transformed_column)

transformedAll <- cbind(non_transformed_column, transformed_column_processed)
rm(customer_table,order_table,product_table,user_features,transformed_column_processed,transformed_column,non_transformed_column)

```


STEP 4: set up training & testing
```{r}
set.seed(1003)

train_rate <- 0.6
training_index <- createDataPartition(transformedAll$is_converted, p = train_rate, list = FALSE, times = 1)

train_data <- transformedAll[training_index,]
test_data <- transformedAll[-training_index,]

train_x <- subset(train_data, select = -c(is_converted))
train_y <- as.factor(apply(subset(train_data, select = c(is_converted)), 2, as.factor))
train_y_categorial <- ifelse(train_y == 1, "YES", "NO")

test_x <- subset(test_data, select = -c(is_converted))
test_y <- as.factor(apply(subset(test_data, select = c(is_converted)), 2, as.factor))
```


STEP 5: Train model

```{r}
registerDoMC(cores=6)

#### Lasso Logistic Regression
model_glm_cv_lasso <- cv.glmnet(data.matrix(train_x),train_y_categorial,alpha = 1,family="binomial",type.measure="auc",parallel=TRUE)

#### Ridge Logistic Regression
model_glm_cv_ridge <- cv.glmnet(data.matrix(train_x),train_y_categorial,alpha = 0,family="binomial",type.measure="auc",parallel=TRUE)

#### Random Forest
rf <- foreach(ntree=rep(200, 6), .combine=combine, .multicombine=TRUE,
              .packages='randomForest') %dopar% {
                randomForest(train_x, train_y, ntree=ntree)
              }
```


STEP 6: Try KNN, Neural Net

```{r}


lasso_predict <- predict(model_glm_cv_lasso, data.matrix(test_x),type='response')

lasso_pred <- prediction(lasso_predict,test_y)
lasso_perf_recall <- performance(lasso_pred,"prec","rec")
lasso_perf_roc <- performance(lasso_pred,"tpr","fpr")
lasso_perf_auc <- performance(lasso_pred,"auc")


repeat_buyers <- subset(
  order_table[order_amount>0,
              .(count=.N),by=customer_id],
  count>1)

orders_sub <- sqldf("SELECT * FROM order_table WHERE customer_id IN (SELECT customer_id FROM repeat_buyers GROUP BY 1)");

orders_sub <- orders_sub[order(orders_sub$customer_id),]
orders_sub$order_rank <- unlist(with(orders_sub, tapply(order_date, customer_id, function(x) rank(x, ties.method= "first"))))
orders_sub$customer_id <- as.character(orders_sub$customer_id)

first_order <- subset(orders_sub[,c("customer_id","order_date","order_rank")],order_rank==1)
second_order <- subset(orders_sub[,c("customer_id","order_date","order_rank")],order_rank==2)

f_to_s <- merge(first_order,second_order,by='customer_id')
f_to_s$f2s_days <- as.numeric(as.Date(as.character(f_to_s$order_date.y),format='%Y%m%d') - as.Date(as.character(f_to_s$order_date.x),format='%Y%m%d'))

orders_sub$order_date <- as.Date(as.character(orders_sub$order_date),format='%Y%m%d')
orders_after2 <- subset(orders_sub,order_rank>=3)
second_order$order_date <- as.Date(as.character(second_order$order_date),format='%Y%m%d')

user_stats <- sqldf("SELECT so.customer_id, SUM(oa.order_amount) AS sum_rev FROM second_order so INNER JOIN orders_after2 oa ON so.customer_id = so.customer_id WHERE oa.order_date - so.order_date <= 183 GROUP BY 1;")
```

